---
title: "LeafletExercise"
author: "A.Casares M."
date: "22 de diciembre de 2016"
output: html_document
---
## The instantaneous world:

With the right tools, presently we can literally have the world
on the tip of our finger, with almost no effort. In this presentation
we'll focus in one of the most awesome of those tools: **Leaflet**, 
here as an add-in to R language.

### Get a mapamundi

```{r, eval=FALSE}
library(leaflet)
mapamundi = ""
mapamundi <- leaflet(mapamundi)
mapamundi <-  addTiles(mapamundi)
mapamundi
```    

```{r, echo=FALSE}
library(leaflet)
mapamundi = ""
mapamundi <- leaflet(mapamundi)
mapamundi <-  addTiles(mapamundi)
mapamundi
```       

You can find on this world's open street map whichever place you like, and go there.

## Put a marker on a special place:

You can add markers to your map one at the time, using the
*addMarkers()* function by specifying the longitude and latitude

```{r, eval=FALSE}
library(leaflet)
my_map <- leaflet()  %>%
  addTiles() %>%
  addMarkers(lat=39.2980803, lng=-76.5898801,
            popup = "Jeff Leek's Office")
my_map
```    

## Your first map:

```{r, echo=FALSE}
library(leaflet)
my_map <- leaflet()  %>%
  addTiles() %>%
  addMarkers(lat=39.2980803, lng=-76.5898801,
  popup = "Jeff Leek's Office")
  # addMarkers(lat=-0.156230, lng=-78.463261,
  #            popup = "Familia Casares Stacey")
my_map
```         

## Adding many Markers:

If you have a dataframe with columns *lat* and *lng* you can pipe
that dataframe into *leaflet()* to add all the points at once.

```{r, eval=FALSE}
set.seed(2016-04-25)
df <- data.frame(lat = runif(20, min = 39.2, max = 39.3),
                 lng = runif(20, min = -76.6, max = -76.5))
                 
df %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers()
```    

```{r, echo=FALSE,warning=FALSE}
set.seed(2016-04-25)
df <- data.frame(lat = runif(20, min = 39.2, max = 39.3),
                 lng = runif(20, min = -76.6, max = -76.5))
                 
df %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers()
```    

## Making Custom Markers

In ocassions, you prefer to use your own design, and not the standard blue markers:   

Also, when adding multiple markers to a map, you may want to add popups for each
marker. You can specify a string of plain text for each popup, or you can
provide HTML which will be rendered inside of each popup.

```{r, eval=FALSE}
hopkinsIcon <- makeIcon(
  iconUrl = "http://brand.jhu.edu/content/uploads/2014/06/university.shield.small_.blue_.png",
  iconWidth = 31*215/230, iconHeight = 31,
  iconAnchorX = 31*215/230/2, iconAnchorY = 16
)

hopkinsLatLong <- data.frame(
  lat = c(39.2973166, 39.3288851, 39.2906617, 39.2970681, 39.2824806),
  lng = c(-76.5929798, -76.6206598, -76.5469683, -76.6150537, -76.6016766))

``` 

```{r, echo=FALSE}
hopkinsIcon <- makeIcon(
  iconUrl = "http://brand.jhu.edu/content/uploads/2014/06/university.shield.small_.blue_.png",
  iconWidth = 31*215/230, iconHeight = 31,
  iconAnchorX = 31*215/230/2, iconAnchorY = 16
)

hopkinsLatLong <- data.frame(
  lat = c(39.2973166, 39.3288851, 39.2906617, 39.2970681, 39.2824806),
  lng = c(-76.5929798, -76.6206598, -76.5469683, -76.6150537, -76.6016766))

```    

## Adding multiple popups:

```{r, eval=FALSE}
hopkinsSites <- c(
  "<a href='http://www.jhsph.edu/'>East Baltimore Campus</a>",
  "<a href='https://apply.jhu.edu/visit/homewood/'>Homewood Campus</a>",
  "<a href='http://www.hopkinsmedicine.org/johns_hopkins_bayview'>Bayview Medical Center</a>",
  "<a href='http://www.peabody.jhu.edu/'>Peabody Institute</a>",
  "<a href='http://carey.jhu.edu/'>Carey Business School</a>"
)

hopkinsLatLong %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(icon = hopkinsIcon, popup = hopkinsSites)
```
```{r, echo=FALSE}
hopkinsSites <- c(
  "<a href='http://www.jhsph.edu/'>East Baltimore Campus</a>",
  "<a href='https://apply.jhu.edu/visit/homewood/'>Homewood Campus</a>",
  "<a href='http://www.hopkinsmedicine.org/johns_hopkins_bayview'>Bayview Medical Center</a>",
  "<a href='http://www.peabody.jhu.edu/'>Peabody Institute</a>",
  "<a href='http://carey.jhu.edu/'>Carey Business School</a>"
)

hopkinsLatLong %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(icon = hopkinsIcon, popup = hopkinsSites)
```

## Mapping clusters

If there are many markers on a zone, and the scale is too short, you cannnot distinguish
the markers from one another. In that case, *leaflet* can cluster them inside a circle, 
with the option *clusterOptions*. Each cluster bears the number of contained markers, and may afterwards be opened via zooming in:

```{r, eval=FALSE}
df <- data.frame(lat = runif(500, min = 39.25, max = 39.35),
                 lng = runif(500, min = -76.65, max = -76.55))
df %>%
  leaflet()  %>%
  addTiles()  %>%
  addMarkers(clusterOptions = markerClusterOptions())
````   

```{r, echo=FALSE}
df <- data.frame(lat = runif(500, min = 39.25, max = 39.35),
                 lng = runif(500, min = -76.65, max = -76.55))
df %>%
  leaflet()  %>%
  addTiles()  %>%
  addMarkers(clusterOptions = markerClusterOptions())
```

## Mapping circle markers:

Instead of adding markers or clusters you can easily add
circle markers using *addCircleMarkers()*.

```{r, eval=FALSE}
df <- data.frame(lat = runif(20, min = 39.25, max = 39.35),
                 lng = runif(20, min = -76.65, max = -76.55))
                 
df %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers()
```     


```{r, echo=FALSE}
df <- data.frame(lat = runif(20, min = 39.25, max = 39.35),
                 lng = runif(20, min = -76.65, max = -76.55))
                 
df %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers()
```     

## Drawing circles
You can draw arbitrary shapes in the maps you create, including circles and squares.
The code below draws a map where the circle on each city is proportional to the
population of that city.

```{r, eval=FALSE}
md_cities <- data.frame(name = c("Baltimore","Frederick","Rockville","Gaithersburg",
                                "Bowie","Hagerstown","Annapolis","College Park",
                                "Salisbury","Laurel"),
                        pop = c(619493, 66169, 62334, 61045, 55232,
                                39890, 38880, 30587, 30484, 25346),
                        lat = c(39.2920592, 39.4143921, 39.084, 39.1434, 39.0068, 39.6418,
                                38.9784, 38.9897, 38.3607, 39.0993),
                        lng = c(-76.6077852,-77.4204875,-77.1528,-77.2014,-76.7791,
                                -77.72,-76.4922,-76.9378,-75.5994,-76.8483)
                        )        
md_cities %>%
  leaflet()  %>%
  addTiles()  %>%
  addCircles(weight = 1, radius = sqrt(md_cities$pop) * 30)

```    

```{r, echo=FALSE, warning=FALSE, comment=""}
md_cities <- data.frame(name = c("Baltimore","Frederick","Rockville","Gaithersburg",
                                "Bowie","Hagerstown","Annapolis","College Park",
                                "Salisbury","Laurel"),
                        pop = c(619493, 66169, 62334, 61045, 55232,
                                39890, 38880, 30587, 30484, 25346),
                        lat = c(39.2920592, 39.4143921, 39.084, 39.1434, 39.0068, 39.6418,
                                38.9784, 38.9897, 38.3607, 39.0993),
                        lng = c(-76.6077852,-77.4204875,-77.1528,-77.2014,-76.7791,
                                -77.72,-76.4922,-76.9378,-75.5994,-76.8483)
                        )        
md_cities %>%
  leaflet()  %>%
  addTiles()  %>%
  addCircles(weight = 1, radius = sqrt(md_cities$pop) * 30)

```     

## Drawing Rectangles:

You can add rectangles - parallel to the window - on *leaflet* maps as well. Just give the coordinates
of two opposed vertices.

```{r, eval=FALSE, warning=FALSE, error=FALSE}
leaflet()  %>%
    addTiles()  %>%
    addRectangles(lat1 = 37.3858, lng1 = -122.0595,
                  lat2 = 37.3890, lng2 = -122.0625)
``` 

```{r, echo=FALSE, warning=FALSE, error=FALSE}
leaflet()  %>%
    addTiles()  %>%
    addRectangles(lat1 = 37.3858, lng1 = -122.0595,
                  lat2 = 37.3890, lng2 = -122.0625)

```    

## Adding legends    

Adding a legend can be useful if you have markers on your map with
different colors.

```{r, eval=FALSE}
df <- data.frame(lat = runif(20, min = 39.25, max = 39.35),
                 lng = runif(20, min = -76.65, max = -76.55),
                 col = sample(c("red","blue","green"),20, replace = TRUE),
                 stringsAsFactors = FALSE)
df %>%
  leaflet()  %>%
  addTiles()  %>%
  addCircleMarkers(color = df$col) %>%
  addLegend(labels = LETTERS[1:3], colors = c("blue","red","green"))
````   

```{r, echo=FALSE, comment=""}
df <- data.frame(lat = runif(20, min = 39.25, max = 39.35),
                 lng = runif(20, min = -76.65, max = -76.55),
                 col = sample(c("red","blue","green"),20, replace = TRUE),
                 stringsAsFactors = FALSE)
df %>%
  leaflet()  %>%
  addTiles()  %>%
  addCircleMarkers(color = df$col) %>%
  addLegend(labels = LETTERS[1:3], colors = c("blue","red","green"))
````   

Isn't it easy and amazing? It's certainly worth of its learning time.

Well, that's all for now, folks!   
